### 2017-04-12 paper/oopsla17-submission/learning.tex
learning edits

Recall that we formulate type error localization as a supervised
classification problem.
%
This means that we assume the existence of a training set
$S = \{\langle \mathbf{v}_i, b_i \rangle\}_{i=1}^{n}$
of sample inputs $\mathbf{v}$ and their corresponding labels $b$.
%
The learning task is to model the probability $Pr(b\ |\ \mathbf{v})$
of a label being assigned to a given input.

There are many learning algorithms to choose from, existing
on a spectrum that balances expressiveness with ease of training (and of
interpreting the learned model).
%
In this section we consider four standard learning algorithms: 
%
(1) logistic regression,
(2) decision trees,
(3) random forests, and
(4) neural networks.
%
A thorough introduction to these techniques can be found in introductory
machine learning textbooks (\eg \ES{TODO: huma/kamalika?}).
%
We will briefly introduce each technique by describing the model it
learns, the training procedure, any hyper-parameters that must be tuned
by a human, and a summary of its advantages and disadvantages.

\ES{FIXME: Wes thinks the entire rest of this section, up to ``4. Evaluation'',
contains much more detail than an OOPSLA audience wants on
introductory/undergraduate machine learning. (Especially since nothing
after this point is a novel contribution we make.) Perhaps we can cut some
space by focusing on the pros and cons of each technique rather than
describing the details? This is a PL paper. }

\paragraph{Logistic Regression}
% The simplest model we investigate is logistic regression, which learns a
% linear function of the features.
% %
% The goal is to learn a set of weights $W$ and biases $b$ such that the
% function
% %
% $$
% Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-W\mathbf{x} - b)}
% $$
% %
% effectively maps feature vectors $\mathbf{x}$ to labels $y$.

@@

The simplest model we investigate is logistic regression, which can be
understood in two steps.
%
First, we compute the \emph{evidence} for a label $b$ as a linear
function of the feature vector $\mathbf{v}$, using a weight vector
$\mathbf{w}$ and a bias $w_0$ that will be learned from the training set.
$$
\mathsf{ev} = \sum_j w_j v_j + w_0
$$
Then we use the logistic function (also called a sigmoid function)
$\sigma(x) = 1 / (1 + e^{-x})$ to compress the evidence term into the
$[0,1]$ interval, which can then be interpreted as the probability that
the label $b$ should be applied.
%
This gives us the final model:
$$
Pr(b\ |\ \mathbf{v}) = \frac{1}{1 + \mathsf{exp}(-\sum_j w_j v_j - w_0)}
$$
% $$
% Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + e^{-\mathsf{ev}}}
% $$

Training a logistic regression entails finding optimal values for the
weights and bias, such that a \emph{cost} function is minimized.
%
In practice, the cost function used for logistic regression is the
\emph{cross entropy}, a measure of the similarity between two
probability distributions, between the predicted labels and the ground
truth.
%
The search for optimal values is typically done via \emph{stochastic
  gradient descent}, an iterative method for optimizing a cost function.
%
Starting with an initial estimate of the weights (\eg a normal
distribution), one repeatedly makes predictions for the training data,
computes the cost function for the predictions, and updates the
parameters according to the gradient of the cost function.
%
This process is repeated until the parameters converge, or until a time
limit is exhausted.
%
Stochastic gradient descent requires a hyper-parameter $\eta$, often
called the \emph{learning rate}, which controls how much the weights and
bias are updated at each iteration.

It has been observed that large weights often coincide with
\emph{overfitting} of the model --- where the model performs well on the
training samples but not on the testing samples.
%
Thus, it is common to add a \emph{regularization} term to the cost
function, a popular choice is the $L_2$ norm of the weights
$\sum_j w_j^2$, which has the effect of penalizing the model for
learning large weights \citep{Park2008-no}.
%
The contribution of the regularization term to the overall cost function
is controlled by another hyper-parameter $\lambda$, often called the
regularization rate.

As a generalized linear model, logistic regression is popular for its
ease of training and of interpreting the resulting model; however, it
can be limited in its applicability.
%
In particular, it may perform poorly when asked to learn an inherently
\emph{nonlinear} function.
%
This limitation can be mitigated by adding transformed or compound
features to the model --- \eg if one knows that the label has a
quadratic relationship with a feature $v$, the transformed feature $v^2$
can be added --- but this typically requires knowledge of the underlying
model.

@@

\paragraph{Decision Trees}
Decision tree algorithms learn a tree of binary predicates over the
features, recursively partitioning the input space until a final
classification can be made.
%
Each node in the tree contains a single predicate of the form
$v_j \leq t$ for some feature $v_j$ and threshold $t$, which determines
whether a given input should proceed down the left or right subtree.
%
Each leaf is labeled with a prediction and the fraction of training
samples that would reach it, the latter quantity can be interpreted as
the decision tree's confidence in its prediction.
\ES{HUMA: is this correct?}

Training a decision tree entails finding both a set of good partitioning
predicates and a good ordering of the predicates.
%
This is usually done in a top-down greedy manner by selecting the
predicate the best partitions the entire training set, and recursively
partitioning the subsets.
%
Standard algorithms for training decision trees include C4.5
\citep{Quinlan1993-de} and CART \citep{Breiman1984-qy}.
%
Overfitting can also be a problem for decision trees; the countermeasure
is usually either halting the learning process before the tree grows too
deep, or pruning the resulting tree.

The advantages of decision trees are that they are relatively easy to
train and that they produce a \emph{white-box} model, making it trivial
to analyze the entire model and explain individual predictions.
%
The disadvantages are that they are prone to overfitting and that the
model can be unstable, \ie small changes to the training data may
produce large changes in the model.
%
Both disadvantages are addressed by random forests.

\paragraph{Random Forests}
One technique for increasing the accuracy of a classifier is to train an
\emph{ensemble} of distinct classifiers and use a majority vote
to make the final prediction.
%
Random forests are an instance of this method, with decision
trees as the underlying classifiers.
%
The training process involves taking $N$ random subsets of the training
data and training a separate decision tree on each subset --- the
training process for the decision trees is often modified slightly to
reduce correlation between trees, by forcing the tree to pick features
from a random subset of all features at each node.
%
In order to make a prediction for an unseen sample, the random forest
asks each decision tree to make a prediction, and then predicts the most
common label.

The diversity of the underlying models tends to make random forests less
susceptible to the overfitting and instability of decision trees, but it
also makes the learned model more difficult to interpret.
%
Furthermore, the number of trees in the forest, $N$, becomes an important
hyper-parameter that must be set manually.

>>>>

Recall that we formulate type error detection at a single expression as a supervised classification problem. This means that there is a training data set $S = \{ (x_i, y_i) \}_{i=1}^n$ of labeled examples, and our goal is to use it to build a classifier or a prediction rule that can predict a label for an input $x$. Because we run the classifier on each expression in order to determine the most likely type error location, we also require the classifier to output a {\em{confidence score}} that indicates how certain it is about its prediction.

%Recall that we formulate type error localization as a supervised
%classification problem.
%
%This means that we assume the existence of a training set
%$S = \{\langle \mathbf{v}_i, b_i \rangle\}_{i=1}^{n}$
%of sample inputs $\mathbf{v}$ and their corresponding labels $b$.
%
%The learning task is to model the probability $Pr(b\ |\ \mathbf{v})$
%of a label being assigned to a given input.

There are many learning algorithms to choose from, existing
on a spectrum that balances expressiveness with ease of training (and of
interpreting the learned model). In this section we consider four standard learning algorithms: (1) logistic regression, (2) decision trees, (3) random forests, and (4) neural networks. A thorough introduction to these techniques can be found in introductory
machine learning textbooks (\eg \cite{FriedmanHastieTibshirani2000}).
% \ES{TODO: huma/kamalika?}).
%

Below we briefly introduce each technique by describing the model it learns, and summarize its advantages and disadvantages. For our application, we are particularly interested in three properties -- expressiveness, interpretability and proneness to overfitting. Expressiveness measures how complex the prediction rule is allowed to be, and interpretability measures how easy it is to explain the cause of prediction to a human. Finally proneness to overfitting measures how easily the rule generalizes to unseen examples that are not in the training set; a rule that is highly prone to overfitting might perform poorly on a test set even though its training performance is very high.

%We will briefly introduce each technique by describing the model it
%learns, the training procedure, any hyper-parameters that must be tuned
%by a human, and a summary of its advantages and disadvantages.

\ES{FIXME: Wes thinks the entire rest of this section, up to ``4. Evaluation'',
contains much more detail than an OOPSLA audience wants on
introductory/undergraduate machine learning. (Especially since nothing
after this point is a novel contribution we make.) Perhaps we can cut some
space by focusing on the pros and cons of each technique rather than
describing the details? This is a PL paper. } \ES{KC: I completely agree. Perhaps we can add a table here. Judging from what we use, the features we need are: (a) expressivity (b) proneness to overfitting and (c) interpretability. We can rate the four classifiers in these three areas.
}

\paragraph{Logistic Regression.} The simplest classifier we investigate is logistic regression. This is a linear model, where the goal is to learn a set of weights $W$ and a bias $b$ that describe the following model for predicting a label $y$ from a feature vector $x$:
\[ \Pr(y = 1 | x) = \frac{1}{1 + e^{-W^{\top} x - b}} \]
The weights $W$ and $b$ are learnt from training data, and the value of $\Pr(y | x)$ naturally leads to a confidence score. Logistic regression is a widely used classification algorithm, which is largely preferred for its simplicity, relative ease of generalization as well as ease of human interpretability. The main limitation is that the prediction rule is constrained to be a linear combination of the features, and hence relatively simple. While this can be somewhat mitigated by adding higher order (quadratic or cubic) features, this often requires thorough domain knowledge.

% The simplest model we investigate is logistic regression, which learns a
% linear function of the features.
% %
% The goal is to learn a set of weights $W$ and biases $b$ such that the
% function
% %
% $$
% Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-W\mathbf{x} - b)}
% $$
% %
% effectively maps feature vectors $\mathbf{x}$ to labels $y$.

@@

%The simplest model we investigate is logistic regression, which can be
%understood in two steps.
%
%First, we compute the \emph{evidence} for a label $b$ as a linear
%function of the feature vector $\mathbf{v}$, using a weight vector
%$\mathbf{w}$ and a bias $w_0$ that will be learned from the training set.
%$$
%\mathsf{ev} = \sum_j w_j v_j + w_0
%$$
%Then we use the logistic function (also called a sigmoid function)
%$\sigma(x) = 1 / (1 + e^{-x})$ to compress the evidence term into the
%$[0,1]$ interval, which can then be interpreted as the probability that
%the label $b$ should be applied.
%
%This gives us the final model:
%$$
%Pr(b\ |\ \mathbf{v}) = \frac{1}{1 + \mathsf{exp}(-\sum_j w_j v_j - w_0)}
%$$
% $$
% Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + e^{-\mathsf{ev}}}
% $$

%Training a logistic regression entails finding optimal values for the
%weights and bias, such that a \emph{cost} function is minimized.
%
%In practice, the cost function used for logistic regression is the
%\emph{cross entropy}, a measure of the similarity between two
%probability distributions, between the predicted labels and the ground
%truth.
%
%The search for optimal values is typically done via \emph{stochastic
%  gradient descent}, an iterative method for optimizing a cost function.
%
%Starting with an initial estimate of the weights (\eg a normal
%distribution), one repeatedly makes predictions for the training data,
%computes the cost function for the predictions, and updates the
%parameters according to the gradient of the cost function.
%
%This process is repeated until the parameters converge, or until a time
%limit is exhausted.
%
%Stochastic gradient descent requires a hyper-parameter $\eta$, often
%called the \emph{learning rate}, which controls how much the weights and
%bias are updated at each iteration.

%It has been observed that large weights often coincide with
%\emph{overfitting} of the model --- where the model performs well on the
%training samples but not on the testing samples.
%
%Thus, it is common to add a \emph{regularization} term to the cost
%function, a popular choice is the $L_2$ norm of the weights
%$\sum_j w_j^2$, which has the effect of penalizing the model for
%learning large weights \citep{Park2008-no}.
%
%The contribution of the regularization term to the overall cost function
%is controlled by another hyper-parameter $\lambda$, often called the
%regularization rate.

%As a generalized linear model, logistic regression is popular for its
%ease of training and of interpreting the resulting model; however, it
%can be limited in its applicability.
%
%In particular, it may perform poorly when asked to learn an inherently
%\emph{nonlinear} function.
%
%This limitation can be mitigated by adding transformed or compound
%features to the model --- \eg if one knows that the label has a
%quadratic relationship with a feature $v$, the transformed feature $v^2$
%can be added --- but this typically requires knowledge of the underlying
%model.

@@

\paragraph{Decision Trees.} Decision tree algorithms learn a tree of binary predicates over the features, recursively partitioning the input space until a final
classification can be made. Each node in the tree contains a single predicate of the form
$v_j \leq t$ for some feature $v_j$ and threshold $t$, which determines
whether a given input should proceed down the left or right subtree.
Each leaf is labeled with a prediction and the fraction of training
samples that would reach it, the latter quantity can be interpreted as
\ES{HUMA: is this correct?}
the decision tree's confidence in its prediction. This leads to a prediction rule
that can be quite complex depending on the input data.

Training a decision tree entails finding both a set of good partitioning and a good ordering of the predicates. This is usually done in a top-down greedy manner, and there are several standard training algorithms such as C4.5
\citep{Quinlan1993-de} and CART \citep{Breiman1984-qy}.

Another advantage of decision trees is their ease of interpretation --  the decision rule is a white-box model that can be readily described to a human, especially when the tree is small. The main limitation is that these trees are often prone to overfitting; while this is somewhat mitigated in practice by pruning the tree, this is still a well-known problem.

%predicates and a good ordering of the predicates.
%
%This is usually done in a top-down greedy manner by selecting the
%predicate the best partitions the entire training set, and recursively
%partitioning the subsets.
%
%Standard algorithms for training decision trees include C4.5
%\citep{Quinlan1993-de} and CART \citep{Breiman1984-qy}.
%
%Overfitting can also be a problem for decision trees; the countermeasure
%is usually either halting the learning process before the tree grows too
%deep, or pruning the resulting tree.

%The advantages of decision trees are that they are relatively easy to
%train and that they produce a \emph{white-box} model, making it trivial
%to analyze the entire model and explain individual predictions.
%
%The disadvantages are that they are prone to overfitting and that the
%model can be unstable, \ie small changes to the training data may
%produce large changes in the model.
%
%Both disadvantages are addressed by random forests.

\paragraph{Random Forests.} Random forests mitigate the overfitting problem by training an {\em{ensemble}} of distinct decision trees and then using a majority vote to make a prediction. The agreement among the distinct trees in the ensemble forms a natural confidence score. Since each classifier in the ensemble is a decision tree, this still allows for complex and highly expressive classifiers. \ES{Huma: is this the confidence score that we are using?}

%One technique for increasing the accuracy of a classifier is to train an
%\emph{ensemble} of distinct classifiers and use a majority vote
%to make the final prediction.
%
%Random forests are an instance of this method, with decision
%trees as the underlying classifiers.
%
The training process involves taking $N$ random subsets of the training
data and training a separate decision tree on each subset --- the
training process for the decision trees is often modified slightly to
reduce correlation between trees, by forcing the tree to pick features
from a random subset of all features at each node.
%
%In order to make a prediction for an unseen sample, the random forest
%asks each decision tree to make a prediction, and then predicts the most
%common label.

The diversity of the underlying models tends to make random forests less
susceptible to the overfitting, but it
also makes the learned model more difficult to interpret. 
%
%Furthermore, the number of trees in the forest, $N$, becomes an important
%hyper-parameter that must be set manually.
